Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae.

### 2020-11-08 conclusion.tex
More work on the paper.

\section{Conclusion}
\label{sec:conclusion}
The algorithm presented in this paper is a first attempt to infer the time evolution of a
combinaison of the mass of the PNS and its radius based on the universal relations found
in PNS asteroseismology. More precisely, we have considered in this paper the ratio
$r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$
oscillation mode in the GW data. We have especially investigated the performance of the algorithm
in the case of an optimally oriented source detected in a singe GW detector. For Advanced LIGO
or Advanced Virgo, the ratio can be reconstructed for a source in the Galaxy. We have shown
that this is true for a wide range of progenitor masses and that the quality of the inference
mainly depends on the signal-to-noise ratio of the signal. For third generation of GW detectors such
as Einstein Telescope and Cosmic Explorer, the $\mbox{}^2g_2$ will be reconstructible for sources
at distances of several kpc. Cosmic Explorer in its stage 2 configuration is obtaining the best performance
for all waveforms considered here thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range.
Among the three configuration of Einstein Telescope, ET-D is providing the best performance,
especially for the waveforms with the highest progenitor mass (25 $\Msol$ and 40 $\Msol$). Comparing
$d_r$ for ET-B and the other third generations projects, it seems that having a good sensitivity
below \unit[200]{Hz} is important for massive mass progenitor signals.

%The model to infer the ratio as function of the frequency is based on 1D {\sc AENUS-ALCAR} simulations.
%We have observed that using the other numerical code {\sc CoCoNut} inputs lead to systematic bias of the
%ratio reconstruction.

This study does not include the realistic case of operating within a network of detectors.
The sources of GWs we have considered here are optimally oriented. The reported distance at which
we can infer the time evolution of $r=M_{\rm PNS}/R_{\rm PNS}^2$ are thus an upper limit that may be lower
by a factor 2--3 on average for a source located anywhere on the sky. We defer a more realistic
simulation implementation for a forthcoming publication.

@@

Finally, this method can be adapted to other PNS oscillation modes, changing few parameters such as
the frequency range of the beginning of the mode and its monotonic raise or descent. Being able to
reconstruct several modes in the same GW signal would allow to infer individually each of the PNS property.

>>>>

% !TEX root = ccsn.\TeX
\section{Conclusions}
\label{sec:conclusion}

The detection of GW from CCSN may help improve our current understanding of the explosion mechanism of massive stars.  
In this paper we have provided estimates of the detectability prospects of CCSN signals using an approach based on GW associated with convective oscillations of PNS. As shown by~\cite{Torres:2019b} buoyancy-driven g-modes are excited in numerical simulations of CCSN and their time-frequency evolution is linked to the physical properties of the compact remnant through universal relations. Such modes are responsible for a significant fraction of the highly stochastic GW emitted after core bounce. The findings reported in this paper suggest that PNS asteroseismology might be within reach of current and third-generation GW detectors.

In our study we have used a set of 1D CCSN simulations to build a model that relates the  evolution of PNS properties with the frequency of the dominant g-mode, namely the $\mbox{}^2g_2$ mode. This relationship is extracted from the GW data using an algorithm developed for this investigation. This algorithm is  a first attempt to infer the time evolution of a particular combination of the PNS mass and radius based on the universal relations found in~\cite{Torres:2019b}. More precisely, we have considered the ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ derived from the observation of the $\mbox{}^2g_2$ oscillation mode in the numerically generated GW data. The  performance of our method has been estimated employing  simulations of 2D CCSN waveforms covering a progenitor mass range between 11 and 40 solar masses and different equations of state.

We have investigated the performance of the algorithm in the case of an optimally oriented source detected in a singe GW detector. The signals of our sample have been injected into 100 Gaussian noise realisations whose PSD follow the spectra of the different GW detectors we have analyzed. For Advanced LIGO and Advanced Virgo, the ratio $r$ can be reconstructed with a good accuracy in the case of a galactic CCSN (i.e.~distances of ${\cal O}(10\, {\rm kpc})$). This holds for a wide range of progenitor masses, the quality of the inference mainly depending on the signal-to-noise ratio of the signal. For third-generation GW detectors such as the Einstein Telescope and the Cosmic Explorer, however, we obtain an order of magnitude improvement, as the $\mbox{}^2g_2$ ratio can be reconstructed for sources at distances of ${\cal O}(100\, {\rm kpc})$. In particular, Cosmic Explorer in its stage 2 configuration yields the best performance for all waveforms we have considered thanks to its excellent sensitivity in the \unit[100-1000]{Hz} range. Among the three configuration of the Einstein Telescope, ET-D provides the best performance, especially for our set waveforms with the higher progenitor masses (25 $\Msol$ and 40 $\Msol$). Comparing the estimated distances for ET-B and the other third-generations detectors, having a good sensitivity below \unit[200]{Hz} seems the most important factor to detect high mass progenitor signals.

In the present study we have assumed that the sources are optimally oriented. The reported distances at which
we can infer the time evolution of $M_{\rm PNS}/R_{\rm PNS}^2$ must thus be regarded as upper limits. Those figures may decrease by a factor 2--3 on average for a source located anywhere in the sky. Furthermore, we have not considered the detectability prospects of  CCSN waveforms in the realistic case in which the  interferometers operated as a detector network. 
We defer an improved implementation of our approach for a forthcoming publication. Finally, we note that the method discussed in this work can be adapted to other PNS oscillation modes, by simply changing a few parameters such as the initial frequency range of the mode and its monotonic raise or descent. Being able to reconstruct several modes in the same GW signal would potentially allow to individually infer the mass and the radius of the PNS.

@@
